---
title: "P8105 Data Science I - Homework 6"
author: "Kevin S.W.   UNI: ksw2137"
date: "`r format(System.time(), %x`"
output: github_document
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}

# global default settings for chunks
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 10, 
                      fig.align = "center",
                      results = "asis"
                      )

# loaded packages; placed here to be able to load global settings
Packages <- c("tidyverse", "dplyr", "modelr", "mgcv")
invisible(lapply(Packages, library, character.only = TRUE))



# global settings for color palettes
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# theme global setting for ggplot
theme_set(theme_minimal() + 
            theme(legend.position = "bottom") +
            theme(plot.title = element_text(hjust = 0.5, size = 12),
                  plot.subtitle = element_text(hjust = 0.5, size = 8))
          )

```

# Problem 1
First load our birthweight data. With 

```{r birthweight}

birthweight_df <- read_csv("./data/Birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate_at(.vars = vars("babysex", "malform", # convert columns in vars() to factors
                         "mrace", "pnumlbw", "pnumsga", "parity",
                         "frace"
                         ),   
            .funs = funs(factor))

# skimr::skim(birthweight_df)                   # remove "#" to activate skimr check 

```

After loading our birthweight data, we obtain a `r nrow(birthweight_df)` by `r ncol(birthweight_df)` dataset that all started as numeric variables. `babysex`, `malform`, `frace`, `mrace` were converted to categorical variables as they should. Count variables with small number of "groups" are also converted to categoricals. These are `parity`, `pnumlbw` and `pnumsga`. `skimr::skim()` reveals no missing variables. 

Variables are:

* Baby-related:
  * `bwt`: baby’s birth weight (g)
  * `blength`: baby’s length at birth (cm)
  * `bhead`: baby’s head circumference at birth (cm)
  * `babysex`: baby’s sex (male = 1, female = 2)
  * `gaweeks`: gestational age in weeks
  * `malform`: presence of malformations that could affect weight (0 = absent, 1 = present)
* Mom-related:
  * `delwt`: mother’s weight at delivery (lbs)
  * `wtgain`: mother’s weight gain during pregnancy (pounds)
  * `ppwt`: mother’s pre-pregnancy weight (pounds)
  * `ppbmi`: mother’s pre-pregnancy BMI
  * `mheigth`: mother’s height (in.)
  * `momage`: mother’s age at delivery (years)
  * `menarche`: mother’s age at menarche (first menstruation; in years)
  * `mrace`: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
  * `parity`: number of live births prior to this pregnancy
  * `pnumlbw`: previous number of low birth weight babies
  * `pnumsga`: number of prior small-for-gestational-age babies
  * `smoken`: average number of cigarettes smoked per day during pregnancy
* Family/father-related:
  * `frace`: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
  * `fincome`: family monthly income (in hundreds, rounded)

### Model Proposal
Given that the mother is the one "doing the most work", it is unlikely that the father's race plays a role so `frace` will not be considered. From the mom-related variables, `menarche` is likely unrelated to birth and if they are, , as well as `parity`. 

```{r}

# proposed model

```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

Compare your model to two others:

* Model 1: length at birth, gestational age as predictors (main effects only)

```{r}

birthweight_df %>% 
  ggplot(aes(x = wtgain, y = bwt, color = menarche)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm")

model1 <- birthweight_df %>% 
  lm(bwt ~ blength + gaweeks, data = .)

```


* Model 2: head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}

model2 <- birthweight_df %>% 
  lm(bwt ~ bhead * blength * babysex, data = .)

```


Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal"

# Problem 2
In this problem we'll be using weather data for Central Park, NY in 2017. Code adopted from homework website. 

```{r CP_2017_load, cache = TRUE}

# loads weather data; code from homework website.
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```



The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

r̂ 2
log(β̂ 0∗β̂ 1)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 0∗β̂ 1).

Optional post-assignment survey